---
title: "Descriptive plots - Intentionality and flexibility in elephant gestural communication"
author: "Vesta Eleuteri"
output:
  rmdformats::material:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
    cards: true
    code_folding: hide
---
  
# Load environment and required packages

```{r load-libraries}
rm(list=ls())

#Load packages
library(ggplot2)
library(scico)
library(here)
library(dplyr)
library(reshape2)
source("Experiment theme.R")

#knitr settings to set Rmd look
library(knitr)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.width = 8,
    fig.height = 6,
    out.width = "100%",
    cache = TRUE
)
```

# Set directory

```{r set-environment}
here::i_am(file.path("Experiment plots.Rmd")) 
```

# Preliminaries

## Frequencies of gesture types by all individuals and by sex

```{r preliminaries}
setwd("/Users/ve/Desktop/PhD/Analysis/Chapter 2 - Experiment/dataframes")
xdata=read.table("exp_sig_df.csv", header=T, sep=",", fill=T, dec=",", stringsAsFactors=T)
gest_data <- xdata %>%
  filter(!Signal_Record %in% c("None", "Unknown", "Trunk-Burp", 
                               "Croaking", "Head-Shake", "Trunk-Fling_Unc", "Trunk-Reach_Unc", 
                               "Lean-Back", "Trunk-Touch-Leg",
                               "Lean-Towards-Ex", "Trunk-Touch-Ear",
                               "Trunk-Touch-TG", "Trunk-Touch-Face")) 
gest_data=droplevels(gest_data) 

#Signal Records frequencies
table_gest=table(gest_data$Signal_Record)
table_gest
table_gest_sex=table(gest_data$Signal_Record, gest_data$Sex)
table_gest_sex
setwd("/Users/ve/Desktop/PhD/Analysis/Chapter 2 - Experiment/results")
write.table(table_gest, file="Table_gest_types.csv", row.names=T, col.names=T, sep="\t")
write.table(table_gest_sex, file="Table_gest_types_sex.csv", row.names=T, col.names=T, sep="\t")
```

# Bar plot for frequency of direction of gestures

```{r bar-plot-direct-gest}
direct_gest_data <- gest_data %>%
  filter(Gest_Direction %in% c("Apple-tray", "Experimenter", "Self")) #keep only gestures with directionality
direct_gest_data=droplevels(direct_gest_data)
table(direct_gest_data$Gest_Direction)

ggplot(direct_gest_data, aes(x = Gest_Direction, fill=Gest_Direction)) +
  geom_bar() +
  labs(title = "Frequency of gesture direction",
       x = "Gesture direction",
       y = "Frequency", fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "managua",
    begin = 0,
    end = 0.8
  )

ggsave(here::here("plots", "descriptives plots", "freq_gest_direction.png"), width = 10, height = 6)
```

# Bar plot for frequency of gesture tokens by all subjects

```{r bar-plot-gest-act-all-subj, fig.width=15}
sub_order <-c("Doma", "Moka", "Mainos", "Laduma", "Detema", "Hwange", "Tatu",
              "Jock", "Pfumo", "Emely", "Naledi", "Jenet", "Jambo", "Coco", "Tendai",
              "Tandiwhe", "Ntombe")
gest_data$Subject <- factor(gest_data$Subject, levels = sub_order)

ggplot(gest_data, aes(x = Subject, fill=Group)) +
  geom_bar() +
  labs(title = "Frequency of gesture tokens by all subjects",
       x = "Subjects",
       y = "Frequency", fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "navia",
    begin = 0.2,
    end = 0.8
  )
ggsave(here::here("plots", "descriptives plots", "freq_gest_tokens_all_subj.png"), width = 20, height = 6)
```

# Stacked bar plot for number of all gesture types by all subjects

```{r stk-bar-plot-gest-typ-all-subj, fig.width=15}
ggplot(gest_data, aes(x = Signal_Record, fill = Subject)) +
  geom_bar() + 
  labs(title = "Frequency of gesture types by all subjects",
       x = "Gesture type",
       y = "Frequency",
       fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "navia", #maybe replace with Roma
    begin = 0,
    end = 1) +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5) #write x label with signal records at 90 degree angle
  )
ggsave(here::here("plots", "descriptives plots", "freq_gest_types_allsubj.png"), width = 20, height = 15)
```

# Stacked bar plot for number of gesture types by at least 2 subjects

```{r stk-bar-plot-gest-typ-min2-subj, fig.width=15}
gest_data_min_2sub <- gest_data %>%
  group_by(Signal_Record) %>%
  filter(n_distinct(Subject) >= 2) %>%
  ungroup() # Optionally ungroup if you need to perform further operations on the full dataset

ggplot(gest_data_min_2sub, aes(x = Signal_Record, fill = Subject)) +
  geom_bar() + 
  labs(title = "Frequency of gesture types by at least two subjects",
       x = "Gesture type",
       y = "Frequency",
       fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "navia", #maybe replace with Roma
    begin = 0,
    end = 1) +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5) #write x label with signal records at 90 degree angle
  )
ggsave(here::here("plots", "descriptives plots", "freq_gest_types_min2_subj.png"), width = 20, height = 15)
```


# Bar plot for frequency of gesture tokens by sexes

```{r bar-plot-gest-act-sex}
ggplot(gest_data, aes(x = Sex, fill=Sex)) +
  geom_bar() +
  labs(title = "Frequency of gesture tokens by sexes",
       x = "Sex",
       y = "Frequency", fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "grayC",
    begin = 0.1,
    end = 0.8)

ggsave(here::here("plots", "descriptives plots", "freq_gest_tokens_sex.png"), width = 10, height = 6)
```

# Bar plot for frequency of gesture tokens by age groups

```{r bar-plot-gest-act-ages}
ggplot(gest_data, aes(x = Age, fill=Age)) +
  geom_bar() +
  labs(title = "Frequency of gesture tokens by age group",
       x = "Age",
       y = "Frequency", fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "grayC",
    begin = 0.2,
    end = 0.8
  )
ggsave(here::here("plots", "descriptives plots", "freq_gest_tokens_age.png"), width = 10, height = 6)
```

# Table and bar plot for frequency of gesture tokens across groups

```{r bar-plot-gest-act-groups-all-subj}
table_gest_group=table(gest_data$Signal_Record, gest_data$Group)
table_gest_group
setwd("/Users/ve/Desktop/PhD/Analysis/Chapter 2 - Experiment/results")
write.table(table_gest_group, file="Table_gest_types_group.csv", row.names=T, col.names=T, sep="\t")

ggplot(gest_data, aes(x = Group, fill=Group)) +
  geom_bar() +
  labs(title = "Frequency of gesture tokens across groups",
       x = "Group",
       y = "Frequency", fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "navia",
    begin = 0.2,
    end = 0.8
  )
ggsave(here::here("plots", "descriptives plots", "freq_gest_tokens_group.png"), width = 10, height = 6)
```

# Stacked bar plot for number of gesture types across groups (including gesture types by all subjects)

```{r stk-bar-plot-gest-typ-group-all-subj, fig.width=15}
ggplot(gest_data, aes(x = Signal_Record, fill = Group)) +
  geom_bar() + 
  labs(title = "(a) Frequency of gesture types by elephant groups",
       x = "Gesture type",
       y = "Frequency",
       fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "navia",
    begin = 0.2,
    end = 0.8) + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5) #write x label with signal records at 90 degree angle
    )
ggsave(here::here("plots", "descriptives plots", "freq_gest_types_group.png"), width = 20, height = 10)
```

# Stacked bar plot for number of gesture types across groups (including gesture types by at least 2 subjects)

```{r stk-bar-plot-gest-typ-group-min2-subj, fig.width=10}
ggplot(gest_data_min_2sub, aes(x = Signal_Record, fill = Group)) +
  geom_bar() + 
  labs(title = "(b) Frequency of gesture types by elephant groups",
       x = "Gesture type",
       y = "Frequency",
       fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "navia",
    begin = 0.2,
    end = 0.8) + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5) #write x label with signal records at 90 degree angle
    )
ggsave(here::here("plots", "descriptives plots", "freq_gest_types_group_min2subj.png"), width = 20, height = 10)
```

# Stacked bar plot for percentage of gesture types across groups (including gesture types by all subjects)

```{r bar-plot-percent-gest-typ-group-all-subj, fig.width=15}
contingency_table <- table(gest_data$Signal_Record, gest_data$Group) #make contingency table

row_sums <- rowSums(contingency_table) #count n in rows
percentage_row <- sweep(contingency_table, 1, row_sums, FUN = "/") * 100 #convert to percentages
percentage_row_df <- as.data.frame(percentage_row)
colnames(percentage_row_df) <- c("Signal_Record", "Group", "Percentage") #add column names
percentage_row_df <- as.data.frame(percentage_row_df)
percentage_row_df$Signal_Record <- factor(percentage_row_df$Signal_Record, 
                                          levels = sort(unique(percentage_row_df$Signal_Record))) #order alphabetically signal records within groups

ggplot(percentage_row_df, aes(x = Signal_Record, y = Percentage, fill = Group)) +
  geom_bar(stat = "identity") + 
  labs(title = "(a) Percentage of gesture types by elephant groups",
       x = "Gesture type",
       y = "Percentage",
       fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "navia",
    begin = 0.2,
    end = 0.8) + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5) #write x label with signal records at 90 degree angle
    )
ggsave(here::here("plots", "descriptives plots", "percent_gest_types_group.png"), width = 20, height = 10)
```

# Stacked bar plot for percentage of gesture types across groups (including gesture types by at least 2subjects)

```{r bar-plot-percent-gest-typ-group-min2-subj, fig.width=15}
contingency_table2 <- table(gest_data_min_2sub$Signal_Record, gest_data_min_2sub$Group) #make contingency table

row_sums2 <- rowSums(contingency_table2) #count n in rows
percentage_row2 <- sweep(contingency_table2, 1, row_sums, FUN = "/") * 100 #convert to percentages
percentage_row_df2 <- as.data.frame(percentage_row2)
library(reshape2)
colnames(percentage_row_df2) <- c("Signal_Record", "Group", "Percentage") #add column names
percentage_row_df2 <- as.data.frame(percentage_row_df2)
percentage_row_df2$Signal_Record <- factor(percentage_row_df2$Signal_Record, 
                                           levels = sort(unique(percentage_row_df2$Signal_Record))) #order alphabetically signal records within groups

ggplot(percentage_row_df2, aes(x = Signal_Record, y = Percentage, fill = Group)) +
  geom_bar(stat = "identity") + 
  labs(title = "(b) Percentage of gesture types by elephant groups",
       x = "Gesture type",
       y = "Percentage",
       fill = "") + theme_exp() +
  scale_fill_scico_d(
    palette = "navia",
    begin = 0.2,
    end = 0.8) + theme(axis.text.x=element_text(angle=90, hjust=1, vjust=.5) #write x label with signal records at 90 degree angle
    )
ggsave(here::here("plots", "descriptives plots", "percent_gest_types_group_min2subj.png"), width = 20, height = 10)
```

# Save working directory

```{r save-directory}
setwd("/Users/ve/Desktop/PhD/Analysis/Chapter 2 - Experiment")
save.image("Experiment plots.Rdata") #Add Rdata
```

